---
title: "SparkRext - SparkR extension for closer to dplyr"
author: Koji MAKIYAMA (@hoxo_m)
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE, error=TRUE)
knitr::knit_hooks$set(document = function(x) {
  SparkR::sparkR.stop()
  x
})
```

## 1. Overview

[Apache Spark](https://spark.apache.org/) is one of the hottest products in data science.  
Spark 1.4.0 has formally adopted **SparkR** package which enables to handle Spark DataFrames on R.(See [this article](http://databricks.com/blog/2015/06/09/announcing-sparkr-r-on-spark.html))

SparkR is very useful and powerful.  
One of the reasons is that SparkR DataFrames present an API similar to **dplyr**.  

For example:

```{r echo=FALSE, cache=FALSE, eval=FALSE}
suppressWarnings(suppressPackageStartupMessages(library(SparkRext)))
sc <- sparkR.init(master="local")
sqlContext <- sparkRSQL.init(sc)
prior_package(SparkR)
```

```{r}
df <- createDataFrame(sqlContext, iris)
df %>%
  select("Sepal_Length", "Species") %>%
  filter(df$Sepal_Length >= 5.5) %>%
  group_by(df$Species) %>%
  summarize(count=n(df$Sepal_Length), mean=mean(df$Sepal_Length)) %>%
  collect  
```

This is very cool. But I have a little discontent.

One of the reasons that dplyr is so much popular is the functions adopts NSE(non-standard evaluation).

```{r}
library(dplyr)
iris %>%
  select(Sepal.Length, Species) %>%
  filter(Sepal.Length >= 5.5) %>%
  group_by(Species) %>%
  summarize(count = n(), mean = mean(Sepal.Length))
```

It's very smart.  
With NSE, you don't need to type quotations or names of DataFrame that the columns belong to.

I have created **SparkRext** package to use NSE version of the functions in SparkR.

```{r echo=FALSE, cache=FALSE, eval=FALSE}
library(SparkRext)
sc <- sparkR.init(master="local")
sqlContext <- sparkRSQL.init(sc)
prior_package(SparkRext)
```

```{r}
library(SparkRext)
df <- createDataFrame(sqlContext, iris)
df %>%
  select(Sepal_Length, Species) %>%
  filter(Sepal_Length >= 5.5) %>%
  group_by(Species) %>%
  summarize(count=n(Sepal_Length), mean=mean(Sepal_Length)) %>%
  collect  
```
