---
title: "SparkRext - SparkR extension for dplyr"
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE, error=TRUE)
```

```{r echo=FALSE}
knitr::knit_hooks$set(document = function(x) {
  SparkR::sparkR.stop()
  x
})
```

```{r echo=FALSE, cache=FALSE, results='hide'}
library(dplyr)
library(SparkR)
library(SparkRext)
sc <- sparkR.init(master="local")
sqlContext <- sparkRSQL.init(sc)
```

This is a fork of the excellent package [SparkRext](https://github.com/hoxo-m/SparkRext), by [@hoxo-m](https://github.com/hoxo-m), which enables users to use [dplyr](https://github.com/saurfang/dplyr) NSE style calls for all data wrangling functions. However it's still impossible to use these functions for distributed Spark DataFrame and local R DataFrame at the same time. This fork enables such use case as shown below.

The motivation is that while SparkR provides a powerful interface to transform distributed DataFrame and practice machine learning algorithms, R still excels in small data world such as [data visualization](#interoperability-between-sparkr-and-dplyr), small data aggregation and etc. 


## Overview

[Apache Spark](https://spark.apache.org/) is one of the hottest products in data science.  
Spark 1.4.0 has formally adopted **SparkR** package which enables to handle Spark DataFrames on R.(See [this article](http://databricks.com/blog/2015/06/09/announcing-sparkr-r-on-spark.html))

SparkR is very useful and powerful.  
One of the reasons is that SparkR DataFrames present an API similar to **dplyr**.  

For example:

```{r, eval=FALSE}
df <- createDataFrame(sqlContext, iris)
df %>%
  select("Sepal_Length", "Species") %>%
  filter(df$Sepal_Length >= 5.5) %>%
  group_by(df$Species) %>%
  summarize(count = n(df$Sepal_Length), mean = mean(df$Sepal_Length)) %>%
  collect  
```

This is very cool. But I have a little discontent.

One of the reasons that dplyr is so much popular is the functions adopts NSE(non-standard evaluation).

```{r}
library(dplyr)
iris %>%
  select(Sepal.Length, Species) %>%
  filter(Sepal.Length >= 5.5) %>%
  group_by(Species) %>%
  summarize(count = n(), mean = mean(Sepal.Length))
```

It's very smart.  
With NSE, you don't need to type quotations or names of DataFrame that the columns belong to.

The package **SparkRext** have been created to make SparkR be closer to dplyr.

```{r echo=FALSE, eval=FALSE}
library(dplyr)
library(SparkR)
library(SparkRext)
```

```{r}
library(SparkRext)
df <- createDataFrame(sqlContext, iris)
df %>%
  select(Sepal_Length, Species) %>%
  filter(Sepal_Length >= 5.5) %>%
  group_by(Species) %>%
  summarize(count = n(Sepal_Length), mean = mean(Sepal_Length)) %>%
  collect  
```

SparkRext redefines the functions of SparkR to enable NSE inputs.  
As a result, the functions will be able to be used in the same way as dplyr.

## How to install

You can install the package from there.

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("saurfang/SparkRext")
```

## 3. Functions

SparkRext redefines six functions on SparkR.

- `filter()`
- `select()`
- `mutate()`
- `arrange()`
- `summarize()`
- `group_by()`

Note we only redefine the generic functions for Spark DataFrame so these functions can still be used with local data frame.

In this section, these funcions are explained.

For illustration, let’s prepare data.

```{r}
library(dplyr)
library(nycflights13)
library(SparkR)

set.seed(123)
data <- sample_n(flights, 10000)

library(SparkRext)

df <- createDataFrame(sqlContext, data.frame(data))
df %>% head
```

### `filter()`

`filter()` is used to extract rows that the conditions specified are satisfied.

```{r}
df %>% filter(month == 12, day == 31) %>% head
```

```{r}
df %>% filter(month == 12 | day == 31) %>% head
```

Note that `filter()` of SparkR cannot accept multiple conditions at once.

### `select()`

`select()` is used to extract columns specified.

```{r}
df %>% select(year, month, day) %>% head
```

Continuous columns can be extracted using a colon `:`.

```{r}
df %>% select(year:day) %>% head
```

You can use the minus sign `-` to extract columns with the exception of columns specified.

```{r}
df %>% select(-year, -month, -day) %>% head
```

You can also extract columns by using column numbers.

```{r}
df %>% select(1, 2, 3) %>% head
```

You can use the select utility functions in dplyr such as `starts_with()`.

```{r}
df %>% select(starts_with("arr")) %>% head
```

All select utility functions is below.

- `starts_with(match, ignore.case = TRUE)`
- `ends_with(match, ignore.case = TRUE)`
- `contains(match, ignore.case = TRUE)`
- `matches(match, ignore.case = TRUE)`
- `num_range(prefix, range, width = NULL)`
- `one_of(...)`
- `everything()`

Note that `select()` of SparkR cannot accept a variety of input like this.

### `mutate()`

`mutate()` is used to add new columns.

```{r}
df %>% mutate(gain = arr_delay - dep_delay, speed = distance / air_time * 60) %>% head
```

Note that `mutate()` of SparkR cannot accept multiple input at once.  
Furthermore, `mutate()` of SparkR cannot also reuse columns added like below.

```{r}
df %>% mutate(gain = arr_delay - dep_delay, gain_per_hour = gain/(air_time/60)) %>% head
```

### `arrange()`

`arrange()` is used to sort rows by columns specified.

```{r}
df %>% arrange(month, day) %>% head
```

It will be sorted in ascending order if you write just column names.  
If you want to sort in descending order, you can use `desc()`.

```{r}
df %>% arrange(month, desc(day)) %>% head
```

You can also sort by values that are transformed from columns.

```{r}
df %>% arrange(abs(dep_delay)) %>% head
```

### `summarize()`

`summarize()` is used to collapse a DataFrame to a single row.

```{r}
df %>% summarize(count = n(year)) %>% collect
```

Typically, `summarize()` is used with `group_by()` to collapse each group to a single row.

As far as I know, you can use the following functions in `summarize()`.

- `n()`
- `n_distinct()`
- `approxCountDistinct()`
- `mean()`
- `first()`
- `last()`

It seems that other aggregate functions are available in Scala (See [docs](http://spark.apache.org/docs/1.4.0/api/scala/index.html#org.apache.spark.sql.functions$)).

Like dplyr, you can use `summarise()` instead of `simmarize()`.

### `group_by()`

`group_by()` is used to describe how to break a DataFrame down into groups of rows.  
Usually it is used with `summarize()` to collapse each group to a single row.

```{r}
df %>% 
  group_by(tailnum) %>%
  summarize(mean_distance = mean(distance)) %>% 
  head
```

You can indicate multiple colmuns.

```{r}
df %>% 
  group_by(year, month, day) %>%
  summarize(count = n(year)) %>% 
  arrange(year, month, day) %>%
  head
```

Unlike dplyr, only `summarize()` can receive the results of `group_by()`.

## How to use

To install SparkR 1.4.0, the next articles may be useful.

- [How to use SparkR within Rstudio?](http://www.r-bloggers.com/how-to-use-sparkr-within-rstudio/)
- [SparkR with Rstudio in Ubuntu 12.04](http://www.r-bloggers.com/sparkr-with-rstudio-in-ubuntu-12-04/)
- [Installing and Starting SparkR Locally on Windows OS and RStudio](http://blog.danielemaasit.com/2015/07/26/installing-and-starting-sparkr-locally-on-windows-8-1-and-rstudio/)

When you can load SparkR package, you will be also able to use SparkRext package.

```{r eval=FALSE}
# Preparation of data
library(dplyr)
library(nycflights13)
set.seed(123)
data <- sample_n(flights, 10000)

# Load library
library(SparkR)
library(SparkRext)

# Create Spark context and SQL context
sc <- sparkR.init(master="local")
sqlContext <- sparkRSQL.init(sc)

# Create DataFrame
df <- createDataFrame(sqlContext, data.frame(data))

# Play with DataFrame
result <- df %>%
  select(year:day, flight, distance) %>%
  group_by(year, month, day) %>%
  summarize(flight_mean = mean(flight), distance_mean = mean(distance)) %>%
  filter(flight_mean >= 2000, distance_mean >= 1000) %>%
  arrange(year, month, day) %>%
  collect

# Print result
head(result)
```

```{r echo=FALSE}
# Play with DataFrame
df %>%
  select(year:day, flight, distance) %>%
  group_by(year, month, day) %>%
  summarize(flight_mean = mean(flight), distance_mean = mean(distance)) %>%
  filter(flight_mean >= 2000, distance_mean >= 1000) %>%
  arrange(year, month, day) %>%
  head
```

## Interoperability between SparkR and dplyr

The appeal of SparkR is operating a large scale dataset with familiar R syntax.
However it would be a shame if we limit ourselves into relying on SparkR for all data manipulation.
For example, R visualization is very powerful and easy to use. You might want to produce a medium size aggregated local data.frame using SparkR and proceed further slice and dice during charting.

For example, we can look average departure delay in a few different dimensions
```{r}
library(ggplot2)

aggDF <- df %>%
  group_by(year, month, carrier, origin, dest) %>%
  summarise(n = n(dep_delay), dep_delay = mean(dep_delay)) %>%
  collect()

aggDF %>%
  group_by(period = as.Date(paste0(year, "-", month, "-01")), carrier) %>%
  summarise(total = n(), dep_delay = weighted.mean(dep_delay, n)) %>%
  # keep period/carrier with more than 20 observations
  filter(total > 20) %>%
  ggplot(aes(period, dep_delay, color = carrier)) +
  geom_line() +
  theme_bw()

aggDF %>%
  group_by(origin, dest) %>%
  summarise(total = n(), dep_delay = weighted.mean(dep_delay, n)) %>%
  filter(total > 30) %>%
  ggplot(aes(origin, dest, fill = dep_delay)) +
  geom_tile(colour = "white") +
  scale_fill_gradient(low = "white", high = "steelblue")
```

As you can see, the seamless transition from large dataframe to small dataframe can be very powerful.
Data science is a not big data or small data endeavor. Having the same set of functions that allow us 
handle both end of the spectrum in the same project can deliver an really enjoyable experience.
